Seismic Hazard Estimation of Northern Iran Using Smoothed Seismicity

### 2016-07-15 introduction.tex
finished making changes requested by cramer

Most previous studies focus on particular locations or are based on specific categorization of seismogenic faults, which are often subjectively interpreted and limited to data availability. We use the smoothed seismicity method introduced by \citet{Frankel1995}. This approach has been used with success in other regions \citep[e.g.][]{Cao1996, Akinci2004, Kalkan2009}, including Iran \citep{Khodaverdian_2016_BSSA}. Its major feature is the treatment of cataloged instrumental and historical earthquakes as individual point sources as opposed to events that are part of a predefined system of faults in a given tectonic setting. While we recognize that probabilistic seismic hazard analysis based on fault systems can be more realistic and accurate, by choosing the smoothed seismicity method we seek to reduce the uncertainty associated with fault mapping, earthquake location and source mechanism definitions, and to cover areas that would otherwise be underestimated or ignored. In our analysis, we assume earthquake hazard in northern Iran to be dominated by the seismic zones of Azerbaijan, Alborz, and Kopeh-Dagh; but for completeness, we also consider the partial contribution of the areas of Zagros and the central-east seismic zones within our domain of interest (see Fig.~\ref{fig:iran}). In addition, for comparison purposes, we obtain results for the the combination of all seismic zones as a uniform region.

In the following sections we describe in greater detail our region of interest, review the associated tectonics and seismicity, and provide additional information about the adopted approach. Significant aspects of our work include the composition of the earthquake catalog, including catalog completeness, the conversion of earthquake magnitudes, the selection of an appropriate attenuation relationship for the region, and the estimation of the Gutenberg-Richter seismicity parameters. Our analysis includes seismic hazard curves and maps for peak ground acceleration (PGA) considering exceedance probabilities of 2 and 10 percent in 50 years for rock site conditions using different magnitude thresholds. We analyze our results in light of equivalent estimations available in the literature and summarize relevant conclusions.

% *********************************************************************************************************************
% *********************************************************************************************************************

% Pieces of text in original introduction that could or should be used later
% --------------------------------------------------------------------------

% \citet{Ghodrati2013} conducted a seismic risk assessment for the city of Tehran using the HAZUS method \citep{FEMA2003}. They represent that these efforts successfully decrease the total mean damage ratio in Tehran from 0.302739 in 1996 to 0.272859 in 2006.

% Having limited knowledge in seismic sources, \citet{Frankel1995} used this approach to map seismic hazards in central and eastern United States. The major features of Frankel's approach are to abandon tectonic seismic zones and to use point sources in seismic hazard analysis. This method is simpler for hazards calculation and is based solely on recorded seismicity history. In the smoothed-seismicity approach we avoid choosing zone boundaries that are sometimes poorly delineated by data and drawn by subjectively merging geological and seismological information. \citet{Masson2006} studied 19 points of the GPS network that have been installed in the framework of French-Iranian cooperation. At some stations (e.g., the ATTA station) they expect to record significant movement, based on the region's historical activity.  Surprisingly, they recorded velocity of $0\  mm/yr$.

% An important aspect of seismic hazard is the proper definition of attenuation parameters. In a recent effort, \citet{Zafarani2014} reviewed multiple ground motion prediction equations available in the literature and concluded that the most appropriate attenuation relationship to be used in northern Iran was \citet{Kallan2004}

% \citet{Nemati2015} studied the most recent 200 years' seismicity in northern Iran. The frequency of shocks vary widely from one mainshock per 6 years (0.17 event/ year) for the Azerbaijan region to 13 earthquakes per 4 years for the Kopeh-Dagh (3.25 event/year) region. Recorded events and a partially quiet period suggest that strong earthquakes must be expected in Alborz within the next decade, which may cause significant damage to northern Iran.

% In this study we use an updated catalog and the recently confirmed attenuation relationship for northern Iran \citep[i.e.,][]{Kalkan2004} in order to conduct seismic hazard analysis from background seismicity. We follow \citet{Frankel1995} approach to calculate hazard from background seismicity for northern Iran. This differs from the traditional approach in which area source zones are drawn around seismicity or tectonic provinces for the calculation of seismic hazard \citep{Cornell1968}.

% Using newly recorded data, \citet{Zafarani2014} defined the most appropriate attenuation relationship for use in the Azerbaijan, Alborz and Kopeh-Dagh regions. In this study we use an updated catalog and the recently confirmed attenuation relationship for northern Iran \citep[i.e.,][]{Kalkan2004} in order to conduct seismic hazard analysis from background seismicity. We follow \citet{Frankel1995} approach to calculate hazard from background seismicity for northern Iran. This differs from the traditional approach in which area source zones are drawn around seismicity or tectonic provinces for the calculation of seismic hazard \citep{Cornell1968}.

% We computed seismic parameters and catalog completeness in these regions. Zagros and Central Iran seismotectonic regions have considerable effect on the Northern part. In order to consider damage to the historical masonry building (which includes most part of big cities and rural areas) we also considered $M_w$ 4.5 as a threshold for structural damage.

% The smoothed-seismicity method simply assumes that patterns of historical earthquakes predict future activity. Since we do not introduce faults into the model, our results show variability due only to the characteristics of seismicity and ground motion.

% *********************************************************************************************************************
% *********************************************************************************************************************

% Naeem's original Introduction
% -----------------------------

% Iran is situated over Himalayan-Alpide seismic belt, which has frequently experienced strong shaking induced by earthquakes. The occurrence of these earthquakes has imposed notable destruction to the buildings and lifelines, and unfortunately, has caused extensive loss of human life. These earthquakes are categorized in various tectonic seismic zones. Different tectonic seismic regions have been defined for Iran in different studies.  Some of these studies defined more detailed division \citep{Nowroozi1976, Tavakoli1999}, and some of them defined one simplified province \citep{Stocklin1968, Takin1972, Berberian1976}. \citet{Mirzaei1998} divided Iran into five tectonic regions, including Azerbaijan-Alborz, Kopeh-Dagh, Zagros, Central-East Iran, and Makran. Regarding the other tectonic seismic regions, \citet{Karimiparidari2013} divided the Azerbaijan-Alborz tectonic seismic region into two regions, which are Azerbaijan and Alborz Mountain Range (hereinafter, Alborz). Azerbaijan, Alborz and Kopeh-Dagh tectonic seismic regions encompass most of northern Iran. Northern Iran has many highly populated cities (e.g., Tehran, Tabriz, and Mashhad), in which many destructive earthquakes have been reported.

% In the last three decades, different seismic hazard analysis studies \citep[e.g.,][]{Tavakoli1999,Ghodrati2003} substantially improved the Iranian Code of Practice for Seismic Resistant Design of Buildings \citep{BHRC2014}. \citet{Ghodrati2013} conducted a seismic risk assessment for the city of Tehran using the HAZUS method \citep{FEMA2003}. They represent that these efforts successfully decrease the total mean damage ratio in Tehran from 0.302739 in 1996 to 0.272859 in 2006.  Even though the results are satisfactory, the continuous improvement of procedures for defining the seismic hazard at regional (national) and local levels is essential for the optimum design of earthquake-resistant structures. In this study we use another model to overcome the aleatory and epistemic uncertainties regarding earthquake location and source mechanism in northern Iran.\\

% In this study we used smoothed seismicity approach \citep{Frankel1995} in order to do seismic hazard analysis. We divided northern Iran in three seismotectonic regions including Azerbaijan, Alborz, and Kopeh Dagh. We also consider seismicity of part of Central-East Iran and Zagros regions. We computed seismic parameters and catalog completeness in these regions. Zagros and Central Iran seismotectonic regions have considerable effect on the Northern part.   We consider the $Mw \  5$ earthquake as a threshold magnitude  for structural damage. In order to consider damage to the historical masonry building (which includes most part of big cities and rural areas) we also considered $Mw \  4.5$ as a threshold for structural damage to these kind of building.

% Recently, northern Iran has been studied in detail from different seismological points of view. \citet{Nemati2015} studied the most recent 200 years' seismicity in northern Iran. The frequency of shocks vary widely from one mainshock per 6 years (0.17 event/ year) for the Azerbaijan region to 13 earthquakes per 4 years for the Kopeh-Dagh (3.25 event/year) region. Recorded events and a partially quiet period suggest that strong earthquakes must be expected in Alborz within the next decade, which may cause significant damage to northern Iran. Using newly recorded data, \citet{Zafarani2014} defined the most appropriate attenuation relationship for use in the Azerbaijan, Alborz and Kopeh-Dagh regions. In this study we use an updated catalog and the recently confirmed attenuation relationship for northern Iran \citep[i.e.,][]{Kalkan2004} in order to conduct seismic hazard analysis from background seismicity. We follow \citet{Frankel1995} approach to calculate hazard from background seismicity for northern Iran. This differs from the traditional approach in which area source zones are drawn around seismicity or tectonic provinces for the calculation of seismic hazard \citep{Cornell1968}.

% Having limited knowledge in seismic sources, \citet{Frankel1995} used this approach to map seismic hazards in central and eastern United States. The major features of Frankel's approach are to abandon tectonic seismic zones and to use point sources in seismic hazard analysis. This method is simpler for hazards calculation and is based solely on recorded seismicity history. In the smoothed-seismicity approach we avoid choosing zone boundaries that are sometimes poorly delineated by data and drawn by subjectively merging geological and seismological information. Even though probabilistic seismic hazard involving fault sources is more realistic and accurate, recognizing sources could present considerable uncertainty. \citet{Masson2006} studied 19 points of the GPS network that have been installed in the framework of French-Iranian cooperation. At some stations (e.g., the ATTA station) they expect to record significant movement, based on the region's historical activity.  Surprisingly, they recorded velocity of $0\  mm/yr$.

% The smoothed-seismicity method simply assumes that patterns of historical earthquakes predict future activity. We generate maps of peak ground acceleration with 2\% and 10\% probabilities in exceedance of 50 years in northern tectonic seismic zones, including Azerbaijan, Alborz, and Kopeh-Dagh and in central-eastern Iran and a small portion of the northern Zagros region. Since we do not introduce faults into the model, our results show variability due only to the characteristics of seismicity and ground motion.

>>>>

Most previous studies focus on particular locations or are based on specific categorization of seismogenic faults, which are often subjectively interpreted and limited to data availability. We use the smoothed seismicity method introduced by \citet{Frankel1995}. This approach has been used with success in other regions \citep[e.g.][]{Cao1996, Akinci2004, Kalkan2009}, including Iran \citep{Khodaverdian_2016_BSSA}. Its major feature is the treatment of cataloged instrumental and historical earthquakes as individual point sources as opposed to events that are part of a predefined system of faults in a given tectonic setting. While we recognize that probabilistic seismic hazard analysis based on fault systems can be more realistic and accurate, by choosing the smoothed seismicity method we seek to reduce the uncertainty associated with fault mapping, earthquake location and source mechanism definitions, and to cover areas that would otherwise be underestimated or ignored. In our analysis, we assume earthquake hazard in northern Iran to be dominated by the seismic zones of Azerbaijan, Alborz, and Kopeh-Dagh, but also consider the contribution of the areas of Zagros and the central-east seismic zones to the hazard calculation within our domain of interest (see Fig.~\ref{fig:iran}). In addition to this five seismic-zone model, we obtain results for the entire region of interest as if it could be considered as a single uniform seismic zone. We used both these models and the combination of them to interpret seismic hazard in northern Iran.

In the following sections we describe our region of interest in greater detail, review its associated tectonics and seismicity, and provide additional information about the adopted approach. Significant aspects of our work include the composition of the earthquake catalog, including catalog completeness and the conversion of earthquake magnitudes, the selection of an appropriate attenuation relationship for the region, and the estimation of the Gutenberg-Richter seismicity parameters. Our analysis includes seismic hazard curves and maps for peak ground acceleration (PGA) considering exceedance probabilities of 2 and 10 percent in 50 years for rock site conditions for events with moment magnitude $M_w /geq 5$. We analyze our results in light of similar estimates available in the literature and summarize the most relevant conclusions.

### 2016-07-15 parameters.tex
finished making changes requested by cramer

We obtain $b$ and $M_{\max}$ for two different models. In the first model we obtain these parameters independently for each one of the five seismic zones influencing the hazard in northern Iran. That is, we obtain distinct values of $b$ and $M_{\max}$ for the main northern seismic zones of Azerbaijan, Alborz, and Kopeh Dagh, as well as for the Zagros and Central-East seismic zones contributing to our region of interest (see Fig.~\ref{fig:iran} for reference). In the second model we consider the whole study area as a uniform seismic zone. Admittedly, this assumption is unrealistic if considered from the point of view that different seismic zones have different seismicity parameters. Computing results for this uniform model is, nonetheless, useful for the analysis of results and comparisons with other studies. Table \ref{tab:params} shows the parameters $b$ and $M_{\max}$ obtained with HA3 for the five seismic zones and the uniform regional model, along with the observed maximum magnitude, $M_{\max}^{\mathrm{obs}}$; and Fig.~\ref{fig:annualp} shows the annual probability of exceedance as a function of earthquake magnitude.

% *********************************************************************************************************************
% OLD
% *********************************************************************************************************************

% For each of the three seismic regions, Gutenberg-Richter parameters and maximum magnitude ($M_{max}$) were calculated using Seismic Hazard Assessment code (HA3) \citep{kijko2004}. The regional maximum magnitude for each region is estimated by the \citet{Kijko1989} method, which is implemented in the HA3 package. For smoothed seismicity areas, $b-value$ is assumed constant. The $a-value$ can vary spatially and is determined by counting earthquakes above $M$ 3.0 in each grid cells.

% \citet{Karimiparidari2013} applied the Maximum Curvature (MAXC) technique \citep{Wyss1999, Wiemer2000} by ZMap \citep{Wiemer2001} to calculate the level of completeness of instrumental part of the catalog.  Following the \citet{Karimiparidari2013}, we assume the catalog is complete for earthquakes with magnitude 4.5, 4.4, 4.5, 4.5, and 4.4 for Azerbaijan, Alborz,  Kopeh-Dagh, Central Iran, and Zagros tectonic seismic regions, respectively. For the uniform model we used 4.5 as a completeness magnitude. In this study we use those magnitudes of completeness as the magnitude threshold in the calculation of the seismicity parameters. We also used the seismicity parameters of \citet{Karimiparidari2013} as the priory information in HA3 code. The updated values are displayed in Table \ref{tab:b_value}.  Annual occurrence rate of the earthquake for each region is shown in Fig.~\ref{fig:annual_m}.

% *********************************************************************************************************************
% OLDER
% *********************************************************************************************************************

% For each of the three seismic regions, Gutenberg-Richter parameters and Max magnitude ($M_{max}$) were calculated using Seismic Hazard Assessment code (HA3) \citep{kijko2004}. The regional maximum magnitude for each region is estimated by the \citet{Kijko1989} method, which is implemented in the HA3 package. For smoothed seismicity areas, $b-value$ is assumed constant. The $a-value$ can vary spatially and is determined by counting earthquakes above $M$ 3.0 in grid cells.\\

% \citet{Karimiparidari2013} applied the Maximum Curvature (MAXC) technique \citep{Wyss1999, Wiemer2000} by ZMap \citep{Wiemer2001} to calculate the level of completeness of instrumental part of the catalog. In this study we use those magnitudes of completeness as the magnitude threshold in the calculation of the seismicity parameters. We also used the seismicity parameters of this study \citep{Karimiparidari2013} as the priory information in HA3 code. The updated values are displayed in Table 2.

% Following the \citet{Karimiparidari2013}, we assume the catalog is complete for earthquakes with magnitude 4.5, 4.4, and 4.5 for Azerbaijan, Alborz and Kopeh-Dagh tectonic seismic regions, respectively. The completeness of each region can be seen easily from the scatter plot of completeness test.

>>>>

We obtain $b$ and $M_{\max}$ for two different models. In the first model we obtain these parameters independently for each one of the five seismic zones influencing the hazard in northern Iran. We call this the \textit{R Model}. In it we obtain distinct values of $b$ and $M_{\max}$ for the main northern seismic zones of Azerbaijan, Alborz, and Kopeh Dagh, as well as for the Zagros and Central-East seismic zones contributing to our region of interest (see Fig.~\ref{fig:iran} for reference). In the second model we consider the whole study area as a uniform seismic zone. We call it the \textit{U Model}. Admittedly, this assumption is somewhat unrealistic if considered from the point of view that different seismic zones have different seismicity parameters. However, as it will be seen in the Results and Analysis section, computing results for this uniform model was useful because it allowed us to find a combination model computation of hazard in the region of interest as a means to reduce epistemic uncertainty. Table \ref{tab:params} shows the parameters $b$ and $M_{\max}$ obtained with HA3 for the five seismic zones in the R model and the corresponding ones for the U model, along with the observed maximum magnitude, $M_{\max}^{\mathrm{obs}}$; and Fig.~\ref{fig:annualp} shows the annual probability of exceedance as a function of earthquake magnitude.
